---
params:
  titleName: ""
  edata: ""
  emeta: ""
  fdata: ""
  omicData: ""
  omicStats: ""
  omicDiag: ""
  pmart_inputs: ""
title: "`r params$titleName`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(dplyr); library(pmartR); library(knitr); library(DT); library(ggplot2);
library(data.table); library(patchwork); library(purrr)
```

```{r Read in Results}
# This data will be passed to the markdown through shiny

# Original edata, emeta, and fdata files 
edata <- params$edata
emeta <- params$emeta
fdata <- params$fdata

# omicData and stats objects 
omicData <- params$omicData # The omic data (e.g. pepData, proData, lipiData, etc.)
omicDataAttr <- attributes(omicData) # The omic data attributes 
omicDiag <- params$omicData # The omic diagnotics for stats
omicDiagAttr <- attributes(omicDiag) # The omic diagnotics attributes 
omicStats <- params$omicStats # The omic stats 
omicStatsAttr <- attributes(omicStats) # The omic stats attributes

# All inputs from tool 
pmart_inputs <- params$pmart_input
pmart_inputs$ReportCode <- TRUE

# Get the file type 
biomolecule <- "RNA-seq"
```

## Upload, Transform, and Format Data 

### *Expression Data*

An `r biomolecule` expression data file called **`r pmart_inputs$file_edata$name`** 
  was uploaded to the PMART application. This file contains `r length(omicData$e_data) - 1`
samples and `r length(unique(omicData$e_data[[omicDataAttr$cnames$edata_cname]]))` 
transcripts designated by the **`r omicDataAttr$cnames$edata_cname`** column. The original scale of the data was counts and remained unchanged. Missing data is not applicable for RNA-seq counts.

Here is the expression data matrix: 
  
```{r Expression Matrix}
datatable(edata, options = list(scrollX = TRUE))
```

`r if (!is.null(omicDataAttr$cnames$emeta_cname)) {paste0("### *Biomolecule Information*\nAn associated biomolecule information file was also uploaded called **", pmart_inputs$file_emeta$name, "** and the molecule identifier column was designated as **", omicDataAttr$cnames$emeta_cname, "**. The file contains ", length(omicData$e_meta) - 1, " columns of biomolecule information.")}`

```{r Biomolecule Information}
if (!is.null(omicDataAttr$cnames$emeta_cname)) {
  datatable(emeta, options = list(scrollX = TRUE))
}
```


`r if (!("Var1" %in% colnames(omicData$f_data))) {paste0("### *Sample Information*\nAn associated sample information file was also uploaded called **", pmart_inputs$file_fdata$name, "**. Trimmed samples names **", ifelse(pmart_inputs$usevizsampnames == "No", "were not", "were"), "** used. The column in the sample information file which indicates sample names was designated as **", omicDataAttr$cnames$fdata_cname, "**.")}`

`r if (!is.null(omicDataAttr$group_DF)) {paste0("*Grouping Information*")}`

```{r Grouping Information: Main Effects and Covariates}
if (!is.null(omicDataAttr$group_DF)) {
  datatable(fdata, options = list(scrollX = TRUE))
}
```

`r if (!is.null(pmart_inputs$pair_id_col)) {paste0("### *Paired Data*\nPaired data was specified. The column with the pair IDs was designated as **", pmart_inputs$pair_id_col, "**, the group identifier column was designated as **", pmart_inputs$pair_group_col, "**, and the value within the group that contains the denominator was determined to be **", pmart_inputs$pair_denom_col, "**. The file contains ", length(unique(omicData$f_data[[pmart_inputs$pair_id_col]])), " unique pairs.")}`

### *Code*

To generate a **`r biomolecule` object** in pmartR for this dataset, use: 
  
```{r Formulate Upload Code, echo = FALSE}
theUploadString <- "No code can be run, as all omicData objects require an e_data and f_data file."

if (!is.null(pmart_inputs$file_fdata$name)) {
  theUploadString <- paste0('omicObj <- as.', class(omicData), '(e_data = read.csv("', pmart_inputs$file_edata$name, '"), edata_cname = "', omicDataAttr$cnames$edata_cname, '", f_data = read.csv("', pmart_inputs$file_fdata$name, '"), fdata_cname = "', omicDataAttr$cnames$fdata_cname, '")')
}

if (!is.null(omicDataAttr$cnames$emeta_cname) & !is.null(pmart_inputs$file_fdata$name)) {
  theUploadString <- theUploadString %>% 
    substr(1, nchar(theUploadString) - 1) %>%
    paste0(', e_meta = read.csv("', pmart_inputs$file_emeta$name, '"), emeta_cname = "',
           omicDataAttr$cnames$emeta_cname, '")')
}
```

``r theUploadString``

`r if(!is.null(pmart_inputs$file_fdata$name)) {"Next, to set group designations, use:"}`

```{r Formulate Group Designation Code, echo = FALSE}
theGroupString <- "No group designation was conducted."

if (!is.null(pmart_inputs$file_fdata$name)) {
  
  # Pull groups 
  gcols <- capture.output({c(pmart_inputs$gcol1, pmart_inputs$gcol2) %>% .[.!= "None"] %>% dput()})
  if (gcols == 'character(0)') {gcols <- "NULL"}
  
  # Pull covariates
  cvcols <- capture.output({c(pmart_inputs$cvcol1, pmart_inputs$cvcol2) %>% .[.!= "None"] %>% dput()})
  if (cvcols == 'character(0)') {cvcols <- "NULL"}
  
  theGroupString <- paste0('omicObj <- group_designation(omicObj, main_effects = ', gcols, ', covariates = ', cvcols, ')')
  
  # Add paired data information 
  if (!is.null(pmart_inputs$pair_id_col)) {
    theGroupString <- theGroupString %>%
      substr(1, nchar(theGroupString) - 1) %>% 
      paste0(', pair_id = "', pmart_inputs$pair_id_col, '", pair_group = "', pmart_inputs$pair_group_col, '", pair_denom = "', pmart_inputs$pair_denom_col, '")')
  }
  
}
```

``r theGroupString``

`r if (length(omicDataAttr$filters) > 0) {"## Filter Data"}`

```{r Filter: Table Summarizing Applied Filters}
#  Put applied filters in a table
    FilterTable <- NULL
    
    if (length(omicDataAttr$filters) > 0) {
      
      # List filter types
      FilterTypes <- lapply(omicDataAttr$filters, function(filter) {filter$type}) %>% unlist()
      
      # Iterate through all the filters for the table, apply different logic for each filter type
      FilterTable <- do.call(rbind, lapply(omicDataAttr$filters, function(filter) {
        
        # Molecule Filter
        if (filter$type == "moleculeFilt") {
          return(c("Filter" = "Molecule Filter", 
                   "pmartR Function" = "molecule_filter",
                   "Type" = "Biomolecule", 
                   "Parameters" = paste0("Min Number Molecules: ", 
                                         filter$threshold), 
                   "Summary" = paste0(filter$filtered %>% length(), 
                                      " biomolecule(s) were filtered."),
                   "First 5 Filtered" = paste0(
                     filter$filtered[1:5][!is.na(filter$filtered[1:5])], 
                     collapse = " & ")))
        } else
          
          # Total Count Filter
          if (filter$type == "totalCountFilt") {
            return(c("Filter" = "Total Count Filter", 
                     "pmartR Function" = "total_count_filter",
                     "Type" = "Biomolecule", 
                     "Parameters" = paste0("Min biomolecule counts: ", 
                                           filter$threshold),
                     "Summary" = paste0(filter$filtered %>% length(), 
                                        " biomolecules(s) were filtered."),
                     "First 5 Filtered" = paste0(
                       filter$filtered[1:5][!is.na(filter$filtered[1:5])], 
                       collapse = " & ")))
          } else 
            
            # RNA Filter
            if (filter$type == "RNAFilt") {
              if(is.null(filter$threshold$size_library)){
                text <- "Non-zero Filter"
                param_text <- "Minimum Non-zero Counts: "
                thresh <- filter$threshold$min_nonzero
              } else {
                text <- "Library Size Filter"
                param_text <- "Minimum Library Size: "
                thresh <- filter$threshold$size_library
              }
              
              return(c("Filter" = text, 
                       "pmartR Function" = "RNA_filter",
                       "Type" = "Sample", 
                       "Parameters" = paste0(param_text, thresh),
                       "Summary" = paste0(filter$filtered %>% length(), 
                                          " sample(s) were filtered."),
                       "First 5 Filtered" = paste0(
                         filter$filtered[1:5][!is.na(filter$filtered[1:5])], 
                         collapse = " & ")))
            } else 
              
              # Custom filter
              if (filter$type == "customFilt") {
                return(c("Filter" = "Custom Filter", 
                         "pmartR Function" = "custom_filter",
                         "Type" = "Sample or Biomolecule", 
                         "Parameters" = "",
                         "Summary" = paste0(
                           filter$filtered$f_data_remove %>% length(), 
                           " sample(s) were filtered. ",
                           filter$filtered$e_data_remove %>% length(), 
                           " biomolecule(s) were filtered. ", 
                           filter$filtered$e_meta_remove %>% length(),
                           " ",
                           get_emeta_cname(omicData),
                           "(s) were filtered."),
                         "First 5 Filtered" = paste0(
                           filter$filtered$f_data_remove[1:5][
                             !is.na(filter$filtered$f_data_remove[1:5])],
                           filter$filtered$e_data_remove[1:5][
                               !is.na(filter$filtered$e_data_remove[1:5])],
                           filter$filtered$e_meta_remove[1:5][
                             !is.na(filter$filtered$e_meta_remove[1:5])],
                           collapse = " & ")))
                
              } else {
                stop(paste0(filter$type, " is not a recognized filter type."))
                }
        })) %>% data.table()
      
      cbind("Order" = 1:nrow(FilterTable), FilterTable) %>% kable()
    }
      
```

`r if ("Molecule Filter" %in% FilterTable$Filter) {paste0("### *Molecule Filter*\nA molecule filter was applied to the data, which removes biomolecule(s), **", omicDataAttr$cnames$edata_cname, "(s)**, not having at least the minimum number of samples (Min Number Molecules) with non-zero counts.")}`

```{r Molecule Filter, echo = FALSE}
if ("Molecule Filter" %in% FilterTable$Filter) {
  
  # Pull filter information
  filterInfo <- omicDataAttr$filters[[which(FilterTypes == "moleculeFilt")[1]]]
  
  # Generate code chunk 
  molFilt_code <- paste0('omicObj <- applyFilt(molecule_filter(omicObj, use_groups = ', filterInfo$method$use_groups, ', use_batch = ', 
                         filterInfo$method$use_batch, '), omicsData = omicObj, min_num = ', filterInfo$threshold, ')')

}
```

`r if ("Molecule Filter" %in% FilterTable$Filter) paste0("**Code:**\n", molFilt_code)`

`r if ("Total Count Filter" %in% FilterTable$Filter) {paste0("### *Total Count Filter*\nA total count filter was applied to the data, which identifies biomolecules, **", omicDataAttr$cnames$edata_cname, "(s)**, with less than the specified number of counts (Min biomolecule counts) across all samples.")}`

```{r Total Count Filter, echo = FALSE}
if ("Total Count Filter" %in% FilterTable$Filter) {
  
  # Pull filter information
  filterInfo <- omicDataAttr$filters[[which(FilterTypes == "totalCountFilt")[1]]]
  
  # Generate code chunk
  tc_code <- paste0('omicObj <- applyFilt(total_count_filter(omicObj), omicsData = omicObj, min_count = ', filterInfo$threshold, ')')

}
```

`r if ("Total Count Filter" %in% FilterTable$Filter) paste0("**Code:**\n", tc_code)`

`r if ("Non-zero Filter" %in% FilterTable$Filter) {paste0("### *Non-zero Filter*\nA non-zero filter was applied to the data, removing sample(s), **", omicDataAttr$cnames$fdata_cname, "(s)**, with library non-zero counts less than the specified threshold (Minimum Non-zero Counts).")}`

```{r RNAFilt nonzero, echo = FALSE}
if ("Non-zero Filter" %in% FilterTable$Filter) {
  
  # Pull filter information
  filterInfo <- omicDataAttr$filters[[which(FilterTypes == "RNAFilt")]]
  if(length(filterInfo) > 1){
    if(is.null(filterInfo[[1]]$threshold$min_nonzero)){
      filterInfo <- filterInfo[[2]]
    } else filterInfo <- filterInfo[[1]]
  }
  
  # Generate code chunk
  nonzero_code <- paste0(
    'omicObj <- applyFilt(RNA_filter(omicData), omicsData = omicObj, min_nonzero = ', 
    filterInfo$threshold$min_nonzero, 
    ')')
  
}
```

`r if ("Non-zero Filter" %in% FilterTable$Filter) paste0("**Code:**\n", nonzero_code)`

`r if ("Library Size Filter" %in% FilterTable$Filter) {paste0("### *Library Size Filter*\nA library size filter was applied to the data, removing sample(s), **", omicDataAttr$cnames$fdata_cname, "(s)**, with total library size less than the specified threshold (Minimum Library Size).")}`

```{r RNAFilt libsize, echo = FALSE}
if ("Library Size Filter" %in% FilterTable$Filter) {
  
  # Pull filter information
  filterInfo <- omicDataAttr$filters[[which(FilterTypes == "RNAFilt")]]
  if(length(filterInfo) > 1){
    if(is.null(filterInfo[[1]]$threshold$size_library)){
      filterInfo <- filterInfo[[2]]
    } else filterInfo <- filterInfo[[1]]
  }
  
  # Generate code chunk
  libsize_code <- paste0(
    'omicObj <- applyFilt(RNA_filter(omicData), omicsData = omicObj, size_library = ', 
    filterInfo$threshold$size_library, ')')
  
}
```

`r if ("Library Size Filter" %in% FilterTable$Filter) paste0("**Code:**\n", libsize_code)`

`r if ("Custom Filter" %in% FilterTable$Filter) {paste0("### *Custom Filter*\n Custom filters can be used to remove samples, biomolecules or  ", get_emeta_cname(omicData)", s. See ?custom_filter to learn how to make and apply custom filters.")}`


`r if (!is.null(omicStats)) {"## Statistics \n\n"}`


```{r Statistics: Text, echo = FALSE}
if (!is.null(omicStats)) {
  ComparisonsText <- paste("The following groups were compared:", paste0(paste(omicStatsAttr$comparisons, sep = ","), 
    "."), "Reported below are the parameters used in the statistical analysis,",
    "followed by the number of siginificant biomolecules for each comparison.")
}
```

`r if (!is.null(omicStats)) {paste(ComparisonsText)}`

```{r Statistics: Table, echo = FALSE}
    if (!is.null(omicStats)) {
      
      # Write the code 
      stat_code <- paste0(
        'diffexp_seq(omicObj, method = "',
        tolower(str_extract(omicStatsAttr$statistical_test, "[A-Za-z]+")),
        '", p_adjust = "', omicStatsAttr$adjustment_method,
        '", p_cutoff = ', omicStatsAttr$pval_thresh, ')')
      
      # Generate a data table of parameters set 
      data.table("Attribute" = c("Test Method", 
                                 "False Discovery Rate Control",
                                 "Significance Threshold"), 
                 "Value" = c(
                   tolower(str_extract(omicStatsAttr$statistical_test, "[A-Za-z]+")), 
                   omicStatsAttr$adjustment_method,
                   omicStatsAttr$pval_thresh)) %>% kable()
    }
```


```{r Statistics: Plot, echo = FALSE}  
if (!is.null(omicStats)) {
  
  # Generate the plot
  plot(omicStats, bw_theme = TRUE)

}
```

`r if (!is.null(omicStats)) {paste0('**Code**: ', stat_code)}`


