---
params:
  titleName: "Metabolite Template File"
  metabData: ""
  metabStats: ""
  pmart_inputs: ""
title: "`r params$titleName`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(dplyr); library(pmartR); library(knitr); library(DT); library(ggplot2);
library(data.table); library(patchwork); library(purrr)
```

```{r Read In Results}
# This data will be passed to the markdown through shiny
metabData <- params$metabData
metabDataAttr <- attributes(metabData)
metabStats <- params$metabStats
metabStatsAttr <- attributes(metabStats)
pmart_inputs <- params$pmart_inputs
```

## Upload Data 

### *Expression Data*

A molecule expression data file called **`r pmart_inputs$file_edata$name`** was uploaded to pmart. 
The column that designates unique molecules was marked as **`r metabDataAttr$cnames$edata_cname`.**
`r if (metabDataAttr$data_info$data_scale_orig == metabDataAttr$data_info$data_scale) {"The original scale of the data was **metabDataAttr$data_info$data_scale_orig** and remained unchanged."} else {paste0("The original scale of the data was **", metabDataAttr$data_info$data_scale_orig, "** and was changed to **", metabDataAttr$data_info$data_scale, "**.")}` 
The value to denote missing data was **`r pmart_inputs$na_symbol`** and the expression data
**`r ifelse(pmart_inputs$normalized_yn == "0", "was not already", "was already")`** normalized. 

`r if (is.null(metabDataAttr$cnames$emeta_cname) == FALSE) {paste0("### *Biomolecule Information*\nAn associated biomolecule information file was also uploaded called **", pmart_inputs$file_emeta$name, "** and the molecule identifier column was designated as **", metabDataAttr$cnames$emeta_cname, "**.")}`

`r if ("Var1" %in% colnames(metabData$f_data) == FALSE) {paste0("### *Sample Information*\nAn associated sample information file was also uploaded called **", pmart_inputs$file_fdata$name, "**. Trimmed samples names **", ifelse(pmart_inputs$usevizsampnames == "No", "were not", "were"), "** used. The column in the sample information file which indicates sample names was designated as **", metabDataAttr$cnames$fdata_cname, "**.")}`

`r if (pmart_inputs$ReportCode) {"### *Example Code*"}`

```{r eval = FALSE, echo = pmart_inputs$ReportCode}
# This is example code of how to generate a metabolite data object in R.

# First, load pmartR
library(pmartR)
library(pmartRdata)
library(tidyverse)

# Then, pull the example data 
data("metab_edata")
data("metab_fdata")

# Generate the metabolomics data object. Note that NA's have been converted to 0's in this step.
mymetabData <- as.metabData(e_data = metab_edata, 
                            f_data = metab_fdata, 
                            edata_cname = "Metabolite", 
                            fdata_cname = "SampleID")
```

`r if (is.null(metabDataAttr$group_DF) == FALSE) {paste0("## Group Samples\n### *Grouping Information*\nThis table summarizes to which group or covariate each column in the Expression Data belongs.")}`

```{r Grouping Information: Main Effects and Covariates}
if (is.null(metabDataAttr$group_DF) == FALSE) {
  metabDataAttr$group_DF %>% kable()
}
```

## Data Summary

### *Summary Table*

The first column in the table below denotes a property of the metabolite data, and the "Data" column
states that property's value.

```{r Data Summary: Main Table}
metabDataSum <- data.frame("Data" = summary(metabData))
metabDataSum %>% kable()
```

### *Missing Value Table*

In the table below, the first column denotes the sample and the second is the number of missing observations.
The third column represents the second as a proportion of the total number of observations
for that sample. 

```{r Data Summary: Missing Value Table}
metabData$e_data %>%
  select(-pmart_inputs$id_col) %>%
  purrr::map(~sum(is.na(.))) %>%
  data.frame() %>%
  t() %>%
  data.frame() %>%
  setNames("Missing Observations") %>%
  mutate(
    "Proportion Missing" = round(`Missing Observations` / nrow(metabData$e_data), 3),
  ) %>% kable()
```

`r if (pmart_inputs$ReportCode && is.null(metabDataAttr$group_DF) == FALSE) {"### *Example Code*"}`

```{r eval = FALSE, echo = pmart_inputs$ReportCode, include = is.null(metabDataAttr$group_DF) == FALSE}
# This is example code of how to transform, group, and get a summarize the metabolite data in R.

# Transform the data
mymetabData <- edata_transform(omicsData = mymetabData, data_scale = "log2")

# Set group designations
mymetabData <- group_designation(omicsData = mymetabData, main_effects = "Condition")

# Summarize results
summary(mymetabData)

# You can also visualize a heatmap or pca
cor_result(mymetabData) %>% plot()
dim_reduction(mymetabData) %>% plot()
```

`r if(length(metabDataAttr$filters) > 0) {paste0("## Filters\nA total of **", length(metabDataAttr$filters), "** filter(s) were applied. See the table below for descriptions of the filters and their order.\n\n### *Summary of Applied Filters*")}`

```{r Filter: Table Summarizing Applied Filters}
#  Put applied filters in a table
FilterTable <- NULL

if (length(metabDataAttr$filters) != 0) {
  
  # Iterate through all the filters for the table, apply different logic for each filter type
  FilterTable <- do.call(rbind, lapply(metabDataAttr$filters, function(filter) {
    
    # Molecule Filter
    if (filter$type == "moleculeFilt") {
      return(c("Filter" = "Molecule Filter", "Type" = "Biomolecule", "Parameters" = 
               paste0("Min Number Molecules: ", filter$threshold), 
               "Summary" = paste0(filter$filtered %>% length(), " biomolecule(s) were filtered."),
               "First 5 Filtered" = paste0(filter$filtered[1:5][!is.na(filter$filtered[1:5])], collapse = " & ")))
    } else
      
    # CV Filter
    if (filter$type == "cvFilt") {
      return(c("Filter" = "CV Filter", "Type" = "Biomolecule", "Parameters" = 
               paste0("Max CV: ", filter$threshold),
               "Summary" = paste0(filter$filtered %>% length(), " biomolecule(s) were filtered."),
               "First 5 Filtered" = paste0(filter$filtered[1:5][!is.na(filter$filtered[1:5])], collapse = " & ")))
    } else
      
    # imd-ANOVA filter
    if (filter$type == "imdanovaFilt") {
      return(c("Filter" = "imd-ANOVA Filter", "Type" = "Biomolecule", "Parameters" = 
               paste0("Min ANOVA: ", filter$threshold$min_nonmiss_anova,
                      " & Min G-Test: ", filter$threshold$min_nonmiss_gtest),
               "Summary" = paste0(filter$filtered %>% length(), " biomolecule(s) were filtered."),
               "First 5 Filtered" = paste0(filter$filtered[1:5][!is.na(filter$filtered[1:5])], collapse = " & ")))
    } else 
      
    # rmd Filter
    if (filter$type == "rmdFilt") {
      return(c("Filter" = "rMD Filter", "Type" = "Sample", "Parameters" = 
                 paste0("P-Value Threshold: ", filter$threshold, " & Metrics Used: ",
                        pmart_inputs$rmd_metrics %>% paste(collapse = ", ")),
               "Summary" = paste0(filter$filtered %>% length(), " sample(s) were filtered."),
               "First 5 Filtered" = paste0(filter$filtered[1:5][!is.na(filter$filtered[1:5])], collapse = " & ")))
    } else 
      
    # Custom filter
    if (filter$type == "customFilt") {
      return(c("Filter" = "Custom Filter", "Type" = "Sample or Biomolecule", "Parameters" = "",
               "Summary" = paste0(filter$filtered$f_data_remove %>% length(), " sample(s) were filtered. ",
                           filter$filtered$e_data_remove %>% length(), " biomolecule(s) were filtered. ", 
                           filter$filtered$e_meta_remove %>% length(), " e_meta variable(s) were filtered."),
               "First 5 Filtered" = paste0(filter$filtered$f_data_remove[1:5][!is.na(filter$filtered$f_data_remove[1:5])], 
                                           filter$filtered$e_data_remove[1:5][!is.na(filter$filtered$e_data_remove[1:5])],
                                           filter$filtered$e_meta_remove[1:5][!is.na(filter$filtered$e_meta_remove[1:5])],
                                           collapse = " & ")))
      
    } else {stop(paste0(filter$type, " is not a recognized filter type."))}
  
  })) %>% data.table()
  
  cbind("Order" = 1:nrow(FilterTable), FilterTable) %>% kable()
  
}
```

`r if ("Molecule Filter" %in% FilterTable$Filter) {paste0("### *Molecule Filter*\nA molecule filter was applied to the data, which removes biomolecule(s), **", metabDataAttr$cnames$edata_cname, "(s)**, not having at least the minimum number of samples (Min Number Molecules).")}`

`r if ("CV Filter" %in% FilterTable$Filter) {paste0("### *CV Filter*\nA coefficient of variation (CV) filter was applied to the data which removes biomolecule(s), **", metabDataAttr$cnames$edata_cname, "(s)**, with a CV greater than the threshold (Max CV).")} `

`r if ("imd-ANOVA Filter" %in% FilterTable$Filter) {paste0("### *imd-ANOVA Filter*\nAn ANOVA filter can be applied to the data which removes biomolecule(s), **", metabDataAttr$cnames$edata_cname, "(s)**, of non-missing values per group (Min ANOVA). Additionaly, an IMD (independence of missing data) filter can be applied to the data, removing biomolecules not having at least a minimum number of non-missing values (Min G-Test) in at least one of the groups.")}`

`r if ("rMD Filter" %in% FilterTable$Filter) {paste0("### *rMD Filter*\nA robust Mahalanobis distance (rMD) filter was applied to the data, removing sample(s), **", metabDataAttr$cnames$fdata_cname, "(s)**, with an associated rMD p-value of less than the threshold (P-Value Threshold). Metrics used to calculate the p-values are also included (Metrics Used).")}`

`r if ("Custom Filter" %in% FilterTable$Filter) {paste0("### *Custom Filter*\n Custom filters can be used to remove samples, biomolecules or proteins.")}`

`r if (pmart_inputs$ReportCode && length(metabDataAttr$filters) > 0) {"### *Example Code*"}`

```{r eval = FALSE, echo = pmart_inputs$ReportCode, include = length(metabDataAttr$filters) > 0}
# This is example code of how to apply and generate filters. 

# Note: Apply filters only if they make sense given the context of the analysis. You can apply any filter with: 
applyFilt(filter, mymetabData, other_parameters)

# and visualize any filter with: 
plot(filter, other_parameters)

# and summarize any filter.
summary(filter, other_parameters)

# Run the biomolecule filter
moleculeFilt <- molecule_filter(omicsData = mymetabData)
plot(moleculeFilt, min_num = 2)

# Run the coefficient of variation filter
cvFilt <- cv_filter(omicsData = mymetabData)
plot(cvFilt, cv_threshold = 80)

# Run the imd-ANOVA filter before running an imd-ANOVA (statistical analysis step)
imdFilt <- imdanova_filter(omicsData = mymetabData)
plot(imdFilt, min_nonmiss_anova = 2)

# Run a robust Mahalanobis distance filter
rmdFilt <- rmd_filter(omicsData = mymetabData, metrics = c("MAD", "Skewness", "Correlation"))
plot(rmdFilt, pvalue_threshold = 0.0001)

# Run a custom filter. This one remove the Infection1 sample. 
custFilt <- custom_filter(omicsData = mymetabData, f_data_remove = "Infection1")
```

`r if (metabDataAttr$data_info$norm_info$is_normalized) {"## Normalization\nMetabolite data **was** normalized.\n\n### *Manual*"}`

```{r Normalization: Manual, Plot}
if (metabDataAttr$data_info$norm_info$is_normalized) {
  
  if (is.null(metabDataAttr$group_DF) == FALSE) {
    
    plot(metabData, order_by = pmart_inputs$gcol1, color_by = pmart_inputs$gcol1, bw_theme = TRUE)
    
  } else {plot(metabData, bw_theme = TRUE)}  
  
}
```

```{r Normalization: Manual, Data Table}
if (is.null(metabDataAttr$data_info$norm_info$norm_type) == FALSE 
    && metabDataAttr$data_info$norm_info$norm_type == "global") {
  
  SubsetParameters <- ifelse(is.null(metabDataAttr$data_info$norm_info$subset_params), "",
    paste(lapply(names(metabDataAttr$data_info$norm_info$subset_params), function(name) {
      paste0(name, ": ", metabDataAttr$data_info$norm_info$subset_params[[name]])}), collapse = " & "))
  
  data.table("Attribute" = c("Subset Function", "Subset Parameters", "Normalization Function"),
             "Value" = c(metabDataAttr$data_info$norm_info$subset_fn, 
                         SubsetParameters, 
                         metabDataAttr$data_info$norm_info$norm_fn)) %>% kable()
}
```

`r if (pmart_inputs$ReportCode && metabDataAttr$data_info$norm_info$is_normalized) {"### *Example Code*"}`

```{r eval = FALSE, echo = pmart_inputs$ReportCode, include = metabDataAttr$data_info$norm_info$is_normalized}
# This is example code of how to conduct the normalization step. 

# Note: There are many normalization parameters to consider. Use ?normalize_global to see them
normalmetabData <- normalize_global(omicsData = mymetabData, subset_fn = "all", norm_fn = "mean")
```

`r if (is.null(metabStats) == FALSE) {"## Statistical Analysis\n The statistical analysis step **was** run."}`

```{r Analysis: Comparison Text}
ComparisonsText <- ""

if (is.null(metabStats) == FALSE) {
  ComparisonsText <- paste("The following groups were compared:", paste0(paste(metabStatsAttr$comparisons, sep = ","), 
    "."), "Reported below are the parameters used in the statistical analysis,",
    "followed by the number of siginificant biomolecules for each comparison.")
}
```

`r ComparisonsText`

```{r Analysis: Input Parameters}
if (is.null(metabStats) == FALSE) {
  data.table("Attribute" = c("Test Method", "Multiple Comparison Adjustment", 
      "Significance Threshold"), "Value" = c(metabStatsAttr$statistical_test, metabStatsAttr$adjustment_method,
      metabStatsAttr$pval_thresh)) %>% kable()
}

if (is.null(metabStats) == FALSE) {
  metabStatsAttr$number_significant[1:4] %>% kable("latex") 
}

if (is.null(metabStats) == FALSE) {
  plot(metabStats, bw_theme = TRUE)
}
```

`r if (pmart_inputs$ReportCode && is.null(metabStats) == FALSE) {"### *Example Code*"}`

```{r eval = FALSE, echo = pmart_inputs$ReportCode, include = is.null(metabStats) == FALSE}
# This is example code of how to conduct the imdanova step

# Note: You should consider whether you want to conduct an anova (default), gtest (missingness), or both. 
# See ?imd_anova for more information
metabStats <- imd_anova(normalmetabData, test_method = "anova")
```






